### 2024-04-02 paper/paper.qmd
Working on introduction and data sections

---
title: "How crisis affect the low-income and high-income economies differently"
subtitle: "Examples from 2008-09 financial crisis and 2020 Covid-19"
author: 
  - Sean Liu 
thanks: "Code and data are available at: https://github.com/Sluuu/Final-Paper.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(readr)
library(dplyr)
library(arrow)
library(dataverse)
library(WDI)

```

# Introduction

```{r}
#| echo: false
#| warning: false
#| message: false

#load data
data_na<- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

## Defining Low-income economies and High-income economies

According to The World Bank (@WDIincome), as of July 2019, using the Gross National Income (GNI) per capita by using the World Bank Atlas method. The Atlas method smooths exchange rate fluctuations using a three-year moving average. The low-income economies are defined as \$1,025 or less, lower middle-income economies are those with between \$1,026 and \$3,995. The upper middle-income economies are those between \$3,996 and \$12,375. Lastly, the high-income economies are those with a GNI per capita \$12,376 or more.

For simplicity, I will group the low-income economies and the lower middle-income economies together as low-income economies. Also, the upper middle-income economies and high-income economies will be grouped as high-income economies. Therefore, a country that has a GNI per capita less than or equal to \$3995 will be categorized as low-income country. A country that has a GNI per capita larger than \$3995 will be categorized as high-income country.

\newpage

# Data

The data is obtained from the World Development Indicators and Other World Bank Data, WDI,  @WDI. The raw data set includes, 17024 observations and 11 variables. As the raw data-set recorded some variables that

- country: It records the country's name.
- year: It records the year of the data being recorded. In our cleaned data set there will only be year 2008, 2009, and 2020.
- inflation: It records the inflation, measured by the consumer price index which reflects the annual percentage change in the cost to the average consumer of acquiring a basket of goods and services that may be fixed or changed at specified intervals, such as yearly.
- gdp_growth: It records the annual percentage growth rate of GDP at market prices. The calculations are based on constant local currency.
- population: It records the total population based on all residents regardless of legal status or citizenship.
- unem_rate: It records the unemployment rate, the unemployment refers to the labours that are not working but is available to work or seeking for employment. 
- lf_par_rate: It records the labor force participation rate, it calculated the proportion of the population who ages 15 and older that are active economically. 
- trade_per: It records the sum of exports and imports of goods and services measured as a share of GDP.
- gni_per_cap: It records the GNI per capita, the unit is converted to U.S. dollars using the World Bank Atlas method. GNI is the sum of value added by all resident producers plus any product taxes. It does not include the value of output plus net receipts of primary income from abroad.

```{r}
#| label: fig-data
#| fig-cap: Relationship between GDP growth versus Inflation
#| echo: false
#| warning: false
#| message: false

data_na |>
  ggplot(mapping = aes(x = gdp_growth, y = inflation)) +
  geom_point() +
  theme_minimal() +
  labs(x = "GDP growth", y = "Inflation")
```

Talk more about it.

# Model

The model is set up by using

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References

### 2024-04-04 paper/paper.qmd
Working on introduction

@@
 
+Globalization improves the connection of different countries of the world.
+Economically,  globalization benefits different countries in multiple ways and being able to trade goods across borders.[@globalization] However, the effects of globalization brings the whole economy on the same boat.
+Consequently, when various global crises occur, they have multiple and different impacts on the world economy.
+
+In this paper, I am interested on how global crises impact each country differently.
+I have chosen the top two global crisis as my study, which are the 2008, 09 global financial crisis and the 2020 COVID-19 pandemic.
 ```{r}
@@
 data_na<- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+#dd_test <- read_parquet(here::here("data/analysis_data/data_na.parquet"))
 ```
@@
 
-According to The World Bank (@WDIincome), as of July 2019, using the Gross National Income (GNI) per capita by using the World Bank Atlas method.
+According to The World Bank [@WDIincome], as of July 2019, using the Gross National Income, GNI, per capita by using the World Bank Atlas method.
 The Atlas method smooths exchange rate fluctuations using a three-year moving average.
@@
 
-- country: It records the country's name.
-- year: It records the year of the data being recorded. In our cleaned data set there will only be year 2008, 2009, and 2020.
-- inflation: It records the inflation, measured by the consumer price index which reflects the annual percentage change in the cost to the average consumer of acquiring a basket of goods and services that may be fixed or changed at specified intervals, such as yearly.
-- gdp_growth: It records the annual percentage growth rate of GDP at market prices. The calculations are based on constant local currency.
-- population: It records the total population based on all residents regardless of legal status or citizenship.
-- unem_rate: It records the unemployment rate, the unemployment refers to the labours that are not working but is available to work or seeking for employment. 
-- lf_par_rate: It records the labor force participation rate, it calculated the proportion of the population who ages 15 and older that are active economically. 
-- trade_per: It records the sum of exports and imports of goods and services measured as a share of GDP.
-- gni_per_cap: It records the GNI per capita, the unit is converted to U.S.
+## Variables 
+- `country`: It records the country's name.
+- `year`: It records the year of the data being recorded. In our cleaned data set there will only be year 2008, 2009, and 2020.
+- `inflation`: It records the inflation, measured by the consumer price index which reflects the annual percentage change in the cost to the average consumer of acquiring a basket of goods and services that may be fixed or changed at specified intervals, such as yearly.
+- `gdp_growth`: It records the annual percentage growth rate of GDP at market prices. The calculations are based on constant local currency.
+- `population`: It records the total population based on all residents regardless of legal status or citizenship.
+- `unem_rate`: It records the unemployment rate, the unemployment refers to the labours that are not working but is available to work or seeking for employment. 
+- `lf_par_rate`: It records the labor force participation rate, it calculated the proportion of the population who ages 15 and older that are active economically. 
+- `trade_per`: It records the sum of exports and imports of goods and services measured as a share of GDP.
+- `gni_per_cap`: It records the GNI per capita, the unit is converted to U.S.
 dollars using the World Bank Atlas method.
@@
 
-Talk more about it.
+In @fig-data, I have presented the relationship between GDP growth and Inflation of each countries from my data-set. Which each dot represents one country.
 
@@
 
-Weaknesses and next steps should also be included.
+If two different time events occurs does low-income countries being impacted differently before, after they become high-income countries?
 

### 2024-04-07 paper/paper.qmd
Update Intro

@@
 title: "How crisis affect the low-income and high-income economies differently"
-subtitle: "Examples from 2008-09 financial crisis and 2020 Covid-19"
+subtitle: "An example from the 2020 Covid-19"
 author: 
@@
 
-In this paper, I am interested on how global crises impact each country differently.
-I have chosen the top two global crisis as my study, which are the 2008, 09 global financial crisis and the 2020 COVID-19 pandemic.
+In this paper, my focus is on the varying impacts of global crises on low and high income countries.
+I have selected the 2020 COVID-19 pandemic as the primary subject of my study.
+This crises have led to varying degrees of economic uncertainty and fluctuations in economic performance worldwide.[@compare] The World Bank indicates that the COVID-19 pandemic has widen the inequality between low and high-income countries, with low-income countries experiencing higher unemployment and inflation rates, along with slower GDP growth.
+These trends align with the expectations set in my study.[@inequality]
 ```{r}
@@
 The raw data set includes, 17024 observations and 11 variables.
-As the raw data-set recorded some variables that
+As the raw data-set recorded some variables have missing values and some repeated values.
+After combining and deleting the extra and missing values, I have 112 observations and 9 variables.
+Which each observations represents a country.
 
@@
 - `country`: It records the country's name.
-- `year`: It records the year of the data being recorded. In our cleaned data set there will only be year 2008, 2009, and 2020.
+- `year`: It records the year of the data being recorded. In our cleaned data set there will only be year 2020.
 - `inflation`: It records the inflation, measured by the consumer price index which reflects the annual percentage change in the cost to the average consumer of acquiring a basket of goods and services that may be fixed or changed at specified intervals, such as yearly.

### 2024-04-07 paper/paper.qmd
Finish intro

@@
 A country that has a GNI per capita larger than \$3995 will be categorized as high-income country.
+Which I have grouped out there are 27 low-income countries in this definition and 85 high-income countries.
 
@@
 As the raw data-set recorded some variables have missing values and some repeated values.
-After combining and deleting the extra and missing values, I have 112 observations and 9 variables.
+After combining and deleting the extra and missing values, I have 112 observations and 10 variables.
 Which each observations represents a country.
@@
 ## Variables 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+#Adding a new variable
+data_na <- data_na %>%
+  mutate(low_income = as.numeric(gni_per_cap <= 3995))
+```
+
 - `country`: It records the country's name.
@@
 It does not include the value of output plus net receipts of primary income from abroad.
+- `low_income`: It is a dummy variable indicating that if a country has a GNI per capita less than or equal to \$3995 it will be 1, else, 0.
 
@@
 
-In @fig-data, I have presented the relationship between GDP growth and Inflation of each countries from my data-set. Which each dot represents one country.
+In @fig-data, I have presented the relationship between GDP growth and Inflation of each countries from my data-set.
+Which each dot represents one country.
+From @fig-data, we could see that countries that have higher/positive gdp growth typically also has a higher inflation rate.
+
+```{r}
+#| label: fig-inflation
+#| fig-cap: Relationship between differenct level of income countries versus Inflation
+#| echo: false
+#| warning: false
+#| message: false
+
+data_na |>
+  ggplot(mapping = aes(x = low_income, y = inflation)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Low-income countries", y = "Inflation")+
+  scale_x_continuous(breaks = c(0,1))
+```
+
+```{r}
+#| label: fig-gdp growth
+#| fig-cap: Relationship between differenct level of income countries versus gdp growth
+#| echo: false
+#| warning: false
+#| message: false
+
+data_na |>
+  ggplot(mapping = aes(x = low_income, y = gdp_growth)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Low-income countries", y = "gdp_growth")+
+  scale_x_continuous(breaks = c(0,1))
+```
+
+```{r}
+data_na |>
+  ggplot(mapping = aes(x = low_income, y = lf_par_rate)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Low-income countries", y = "Inflation")+
+  scale_x_continuous(breaks = c(0,1))
+```
+
+\newpage
 

### 2024-04-07 paper/paper.qmd
Finsih Abstract

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper examines the impacts of global crises focusing on the 2020 COVID-19 pandemic's effects on low and high-income countries.
+It highlights how globalization links economies, leading to shared challenges during worldwide disturbances.
+The analysis reveals that low-income countries have suffered more acutely, experiencing higher unemployment, inflation, and slower GDP growth."
 format: pdf

### 2024-04-08 scripts/01-download_data.R
Fixing typo

@@ -33,3 +33,4 @@ world_bank_data <-WDI(
 
 #### Save Data ####
 write_csv(world_bank_data, "data/raw_data/raw_data.csv")
+

### 2024-04-08 scripts/02-data_cleaning.R

@@ -24,15 +24,18 @@ wb_data <-
     unem_rate = SL.UEM.TOTL.NE.ZS,
     lf_par_rate = SL.TLF.CACT.NE.ZS,
     trade_per = NE.TRD.GNFS.ZS, 
-    gdp_per_cap = NY.GDP.PCAP.CD
+    gni_per_cap = NY.GNP.PCAP.CD
   ) |>
-  select(country, year, inflation, gdp_growth, population, unem_rate, lf_par_rate, trade_per, gdp_per_cap) |>
-  filter(year %in% c(2020, 2008, 2009))
+  select(country, year, inflation, gdp_growth, population, unem_rate, lf_par_rate, trade_per, gni_per_cap) |>
+  filter(year %in% c(2020))
+
+
+
 data_na <- wb_data |> drop_na()
 
 
-
 #### Save data ####
-write_parquet(data_na, "data/analysis_data/data_na.parquet")
+
+#write_parquet(data_na, "data/analysis_data/data_na.parquet")
 write_csv(data_na, "data/analysis_data/analysis_data.csv")
 

### 2024-04-09 paper/paper.qmd
Working on Data section

@@
 library(WDI)
+library(rstanarm)
 
@@
 
+\newpage
+
 ```{r}
@@
 ```
+In @fig-inflation, I have shown the inflation rate for the low-income and the high-income countries.
+On the x-axis, when x is equal to 0, it means that the county is a high-income country and it equals to 1 when the country is a low-income country.
+From @fig-inflation, we could see that high income countries has lower and more stabilized inflation rates.
+In low-income countries the inflation rate varies across each country.
+
+\newpage
 
 ```{r}
-#| label: fig-gdp growth
+#| label: fig-gdp-growth
 #| fig-cap: Relationship between differenct level of income countries versus gdp growth
@@
 ```
+In @fig-gdp-growth, I have shown the gdp growth rate for the low-income and the high-income countries.
+On the x-axis, when x is equal to 0, it means that the county is a high-income country and it equals to 1 when the country is a low-income country.
+From @fig-gdp-growth, we could see that high income countries have similar gdp growth rates.
+In low-income countries the gdp growth rates varies across each country.
+
+\newpage
 
 ```{r}
+#| label: fig-lf
+#| fig-cap: Relationship between differenct level of income countries versus labour force participation rate
+#| echo: false
+#| warning: false
+#| message: false
 data_na |>
@@
   theme_minimal() +
-  labs(x = "Low-income countries", y = "Inflation")+
+  labs(x = "Low-income countries", y = "labour force rate")+
   scale_x_continuous(breaks = c(0,1))
 ```
+
+From the data presented in @fig-inflation, @fig-gdp-growth, and @fig-lf, I observed that high-income countries shared similar economic performance trends.
+Despite not necessarily excelling in all economic indicators than the low-income countries, however, these countries generally maintain a more stable economic condition, with only a few exceptions or outliers noted.
+
+ A more detailed relationship between each variable could be found in @fig-relationship.
 
@@
 The model is set up by using
+
+```{r}
+
+lowincome <- data_na %>% filter(low_income == 1)
+highincome <- data_na %>% filter(low_income == 0)
+lowmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 
+             family = gaussian(link = "identity"), data = lowincome)
+
+summary(lowmodel)
+
+```
+
+```{r}
+highmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 
+             family = gaussian(link = "identity"), data = highincome)
+
+summary(highmodel)
+```
 
@@
 
-# Appendix {-}
+# Additional data details
+```{r}
+#| label: fig-relationship 
+#| fig-cap: Scatterplot matrix of each variables
+#| echo: false
+#| warning: false
+#| message: false
 
-# Additional data details
+pairs(~inflation + gdp_growth + population + unem_rate + lf_par_rate + trade_per + gni_per_cap, data = data_na)
+```
 

### 2024-04-09 paper/paper.qmd
Working on Models

@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 highmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 

### 2024-04-13 paper/paper.qmd
Adding year 2021 into the study

@@
 title: "How crisis affect the low-income and high-income economies differently"
-subtitle: "An example from the 2020 Covid-19"
+subtitle: "An example from the 2020/2021 Covid-19"
 author: 
@@
 date-format: long
-abstract: "This paper examines the impacts of global crises focusing on the 2020 COVID-19 pandemic's effects on low and high-income countries.
+abstract: "This paper examines the impacts of global crises focusing on the 2020 and 2021 COVID-19 pandemic's effects on low and high-income countries.
 It highlights how globalization links economies, leading to shared challenges during worldwide disturbances.
@@
 In this paper, my focus is on the varying impacts of global crises on low and high income countries.
-I have selected the 2020 COVID-19 pandemic as the primary subject of my study.
+I have selected the 2020 and 2021 COVID-19 pandemic as the primary subject of my study.
 This crises have led to varying degrees of economic uncertainty and fluctuations in economic performance worldwide.[@compare] The World Bank indicates that the COVID-19 pandemic has widen the inequality between low and high-income countries, with low-income countries experiencing higher unemployment and inflation rates, along with slower GDP growth.
@@
 As the raw data-set recorded some variables have missing values and some repeated values.
-After combining and deleting the extra and missing values, I have 112 observations and 10 variables.
+After combining and deleting the extra and missing values, In the year of 2020, I have 112 observations and in the year of 2021 I have 117 variables which both years with the sharing 10 variables.
 Which each observations represents a country.
@@
 data_na |>
-  ggplot(mapping = aes(x = gdp_growth, y = inflation)) +
+  ggplot(mapping = aes(x = gdp_growth, y = inflation, color = as.factor(year))) +
   geom_point() +
   theme_minimal() +
-  labs(x = "GDP growth", y = "Inflation")
+  labs(x = "GDP Growth", y = "Inflation", color = "Years")
 ```
@@
 data_na |>
-  ggplot(mapping = aes(x = low_income, y = inflation)) +
+  ggplot(mapping = aes(x = low_income, y = inflation, color = as.factor(year))) +
   geom_point() +
   theme_minimal() +
-  labs(x = "Low-income countries", y = "Inflation")+
+  labs(x = "Low-income countries", y = "Inflation", color = "Years")+
   scale_x_continuous(breaks = c(0,1))
 ```
-In @fig-inflation, I have shown the inflation rate for the low-income and the high-income countries.
+In @fig-inflation, I have shown the inflation rate for the low-income and the high-income countries in 2020 and 2021.
 On the x-axis, when x is equal to 0, it means that the county is a high-income country and it equals to 1 when the country is a low-income country.
-From @fig-inflation, we could see that high income countries has lower and more stabilized inflation rates.
-In low-income countries the inflation rate varies across each country.
+From @fig-inflation, we could also see that high income countries has lower and more stabilized inflation rates in either year 2020 and year 2021.
+Which in low-income countries the inflation rate varies across each country and there are also more outliers in year 2021 compare with 2020.
 
@@
 data_na |>
-  ggplot(mapping = aes(x = low_income, y = gdp_growth)) +
+  ggplot(mapping = aes(x = low_income, y = gdp_growth, color = as.factor(year))) +
   geom_point() +
   theme_minimal() +
-  labs(x = "Low-income countries", y = "gdp_growth")+
+  labs(x = "Low-income countries", y = "gdp_growth", color = "Years")+
   scale_x_continuous(breaks = c(0,1))
@@
 On the x-axis, when x is equal to 0, it means that the county is a high-income country and it equals to 1 when the country is a low-income country.
-From @fig-gdp-growth, we could see that high income countries have similar gdp growth rates.
-In low-income countries the gdp growth rates varies across each country.
+From @fig-gdp-growth, we could see that in both 2020 and 2021, each year the high income countries have similar gdp growth rates.
+Which in low-income countries the gdp growth rates more varies across each country.
+Lastly in both low and high income situations, the overall gdp growth in 2021 are higher than in 2020.
 
@@
 data_na |>
-  ggplot(mapping = aes(x = low_income, y = lf_par_rate)) +
+  ggplot(mapping = aes(x = low_income, y = lf_par_rate, color = as.factor(year))) +
   geom_point() +
   theme_minimal() +
-  labs(x = "Low-income countries", y = "labour force rate")+
+  labs(x = "Low-income countries", y = "labour force rate", color = "Years")+
   scale_x_continuous(breaks = c(0,1))

### 2024-04-13 scripts/02-data_cleaning.R

@@ -27,7 +27,7 @@ wb_data <-
     gni_per_cap = NY.GNP.PCAP.CD
   ) |>
   select(country, year, inflation, gdp_growth, population, unem_rate, lf_par_rate, trade_per, gni_per_cap) |>
-  filter(year %in% c(2020))
+  filter(year %in% c(2020, 2021))
 
 
 

### 2024-04-14 paper/paper.qmd
Update Reference

@@
 A country that has a GNI per capita larger than \$3995 will be categorized as high-income country.
-Which I have grouped out there are 27 low-income countries in this definition and 85 high-income countries.
+Which I have grouped out in 2020 there are 27 low-income countries in this definition and 85 high-income countries.
+In 2021, there are
 
@@
 From @fig-data, we could see that countries that have higher/positive gdp growth typically also has a higher inflation rate.
+Also in 2021, the gdp growth for each country are typically higher than in 2020 with having similar inflation rate.
 
@@
 data_na |>
-  ggplot(mapping = aes(x = low_income, y = lf_par_rate, color = as.factor(year))) +
+  ggplot(mapping = aes(x = low_income, y = unem_rate, color = as.factor(year))) +
   geom_point() +
@@
 
-The model is set up by using
+The model is set up by using R [@citeR] with the help of the following packages, `tidyverse` package of @tidyverse, `readr` package of @readr, `dplyr` package of @dplyr, `dataverse` package of @dataverse
 

### 2024-04-15 paper/paper.qmd
Update Model Section

@@
 library(rstanarm)
-
+library(broom)
+library(stargazer)
 ```
@@
 ```{r}
-#| label: fig-lf
+#| label: fig-ur
 #| fig-cap: Relationship between differenct level of income countries versus labour force participation rate
@@
   theme_minimal() +
-  labs(x = "Low-income countries", y = "labour force rate", color = "Years")+
+  labs(x = "Low-income countries", y = "Unemployment Rate", color = "Years")+
   scale_x_continuous(breaks = c(0,1))
@@
 
-From the data presented in @fig-inflation, @fig-gdp-growth, and @fig-lf, I observed that high-income countries shared similar economic performance trends.
+From @fig-ur, the low-income countries seems to have lower unemployment rate than in high-income countries.
+Also, in both 2020 and 2021, the unemployment rates for both low and high income countries have similar results.
+
+Lastly, the data presented in @fig-inflation, @fig-gdp-growth, and @fig-lf, I observed that high-income countries shared similar economic performance trends.
 Despite not necessarily excelling in all economic indicators than the low-income countries, however, these countries generally maintain a more stable economic condition, with only a few exceptions or outliers noted.
+Also the overall economic performance in 2021 are better than in 2020.
 
@@
 
-The model is set up by using R [@citeR] with the help of the following packages, `tidyverse` package of @tidyverse, `readr` package of @readr, `dplyr` package of @dplyr, `dataverse` package of @dataverse
+The model is set up by using R [@citeR] with the help of the following packages, `tidyverse` package of @tidyverse, `readr` package of @readr, `dplyr` package of @dplyr, `dataverse` package of @dataverse, `stargazer` and package of  @stargazer, and `broom` package of @broom.
+Given that countries are classified as low or high income based on their GNI per capita, `gni_per_cap`, I will use this as the dependent variable in a multiple linear regression model to analyze the impact of various factors on `gni_per_cap` across these groups.
+I will explore how different variables affect `gni_per_cap` in both low and high-income countries.
+Specifically, I anticipate that low-income countries will experience larger negative effect from unemployment and inflation rates, as well as experience negative GDP growth.
+This analysis aims to provide insights into the economic challenges faced by these countries and inform potential policy interventions.
+
+\newpage
 
@@
 
-summary(lowmodel)
+stargazer(lowmodel, type = "text",
+          title = "Summary of Low Income Model",
+          align = TRUE,
+          out = "model_summary.txt")
 
 ```
+From the Summary of Low Income Model above, we could see the following situations:
+
+The GDP growth (`gdp_growth`) coefficient is -25.093 with a standard error of 21.917.
+This suggests there is a negative relationship between GDP growth and GNI per capita, though it is not statistically significant as its high p-value.
+Then the Unemployment rate (`unem_rate`) coefficient is -23.374 with a standard error of 34.236.
+This indicates a negative relationship between the unemployment rate and GNI per capita, which is also not statistically significant.
+Lastly the Inflation (`inflation`) coefficient is -7.082 with a standard error of 8.800, indicating a negative relationship between inflation and GNI per capita, though this relationship is not statistically significant.
+
+The Labor force participation rate (`lf_par_rate`) coefficient is -52.818 with a standard error of 13.104, and it is significant at the p<0.01 level.
+This significant negative relationship implies that higher labor force participation is associated with lower GNI per capita among the low-income countries in this analysis, which might suggest underemployment or low-quality jobs prevalent in these economies.
+The Trade percentage of GDP (`trade_per`) coefficient is 11.697 with a standard error of 3.992, significant at the p<0.01 level.
+This significant positive relationship suggests that a higher trade percentage of GDP is associated with higher GNI per capita, indicating the beneficial impact of trade on national income in low-income countries.
+
+The Population (`population`) coefficient is different than the other variables.
+As its coefficient is extremely small and not statistically significant, suggesting that population size, has no impact on GNI per capita in this model.
+
+Constant: The intercept is 5,120.936 with a standard error of 926.753, significant at the p<0.01 level.
+This value represents the estimated GNI per capita when all other predictors are zero, which is a theoretical baseline for the model.
+
+Model Fit and Quality
+Observations: There are 53 observations in this model, providing a somewhat small but possibly adequate sample size for initial explorations in economic data.
+
+Log Likelihood: The log likelihood value is -432.942. While in isolation this number doesn't tell much, it helps in model comparisons, particularly when comparing nested models.
+
+Akaike Information Criterion (AIC): The AIC is 879.883. A lower AIC indicates a better model fit relative to models with higher AIC values, assuming the same data. It is used for model selection.
 
@@
 
-summary(highmodel)
+stargazer(highmodel, type = "text",
+          title = "Summary of High Income Model",
+          align = TRUE,
+          out = "model_summary.txt")
 ```
+The high income-countries have an overall better performance than the low-income countries: 
+The GDP Growth (`gdp_growth`) coefficient is 567.121 with a standard error of 244.367, significant at the p<0.05 level.
+This suggests a positive and statistically significant relationship between GDP growth and GNI per capita in high-income countries, indicating that economic growth positively impacts income levels.
+Then the Unemployment Rate (`unem_rate`) coefficient is -849.519 with a standard error of 350.854, significant at the p<0.05 level.
+This indicates a significant negative impact of higher unemployment rates on GNI per capita, suggesting that unemployment is a critical economic issue affecting income in high-income countries.
+The Labor Force Participation Rate (`lf_par_rate`) coefficient is 433.735 with a standard error of 229.487, significant at the p<0.1 level.
+This shows a positive relationship between labor force participation and GNI per capita, though the relationship is less statistically significant.
+The Trade Percentage of GDP (`trade_per`) coefficient is 59.875 with a standard error of 22.554, significant at the p<0.01 level.
+This indicates a robust positive relationship, suggesting that trade openness contributes significantly to GNI per capita in high-income countries.
+Lastly, the Inflation (`inflation`) coefficient is -2,799.239 with a standard error of 600.123, significant at the p<0.01 level.
+This suggests a strong negative impact of inflation on GNI per capita, indicating that higher inflation rates are detrimental to income levels.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Same with the low-income situation, the Population (`population`) coefficient is extremely small and statistically insignificant, implying no clear impact of population size on GNI per capita within this model.
+
+Constant: The intercept is 7,822.731 with a substantial standard error of 16,169.640.
+While not statistically significant, this value indicates the baseline GNI per capita when all predictors are zero, though this scenario is mostly theoretical given the nature of the predictors.
 
@@
 \end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 

### 2024-04-16 paper/paper.qmd
Working on Model
Update Model adjustment and mordel set up

@@
 
-Constant: The intercept is 5,120.936 with a standard error of 926.753, significant at the p<0.01 level.
-This value represents the estimated GNI per capita when all other predictors are zero, which is a theoretical baseline for the model.
-
-Model Fit and Quality
-Observations: There are 53 observations in this model, providing a somewhat small but possibly adequate sample size for initial explorations in economic data.
-
-Log Likelihood: The log likelihood value is -432.942. While in isolation this number doesn't tell much, it helps in model comparisons, particularly when comparing nested models.
-
-Akaike Information Criterion (AIC): The AIC is 879.883. A lower AIC indicates a better model fit relative to models with higher AIC values, assuming the same data. It is used for model selection.
-
 ```{r}
@@
 
-Constant: The intercept is 7,822.731 with a substantial standard error of 16,169.640.
-While not statistically significant, this value indicates the baseline GNI per capita when all predictors are zero, though this scenario is mostly theoretical given the nature of the predictors.
-
 ## Model set-up
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+To consider the effect of the 6 economic indicators on the level on low and high income countries' GNI per capita, I used the multiple linear regression to approach my data. The formula is as follows:
 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+\text{gni\_per\_cap} &= \alpha + \beta_1 \times \text{population}_i + \beta_2 \times \text{unem\_rate}_i + \beta_3 \times \text{lf\_par\_rate}_i \\
+&\quad + \beta_4 \times \text{trade\_per}_i + \beta_5 \times \text{inflation}_i, \\
+\alpha &\sim \text{Normal}(0, 2.5), \\
+\beta_1 &\sim \text{Normal}(0, 2.5), \\
+\beta_2 &\sim \text{Normal}(0, 2.5), \\
+\beta_3 &\sim \text{Normal}(0, 2.5), \\
+\beta_4 &\sim \text{Normal}(0, 2.5), \\
+\beta_5 &\sim \text{Normal}(0, 2.5)
 \end{align}
 
-### Model justification
+### Model Adjustment
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+For a better performance and interpretation of the model, I have applied the log transformation to my model.
+The log transformation is to reduce the skewness in variables that have a skewed distribution, such as `gni_per_cpa`, `gdp_growth`, `population`, etc.
+By transforming these variables, the distribution is more symmetric, which is closer to the normal distribution.
+The Log transformations also help stabilize the variance across different levels of input variables, reduces homoscedasticity.
+The log transformation variables are the follows:
+Log of `gni_per_cap` The dependent variable's log transformation can help in normalizing the distribution and making relationships more linear.
+The log of `gdp_growth` and `population` shared similar reasons.
+As Economic growth rates and population can vary widely, and the impact on GNI per capita might scale logarithmically rather than linearly.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+log_lowmodel <- glm(log(gni_per_cap) ~ log(gdp_growth + 1) + log(population + 1) + unem_rate + lf_par_rate + trade_per + inflation, family = gaussian(link = "identity"), data = lowincome)
+
+stargazer(log_lowmodel, type = "text",
+          title = "Summary of Low Income Model",
+          align = TRUE,
+          out = "model_summary.txt")
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+log_highmodel <- glm(log(gni_per_cap) ~ log(gdp_growth + 1) + log(population + 1) + unem_rate + lf_par_rate + trade_per + inflation,family = gaussian(link = "identity"), data = highincome)
+
+stargazer(log_highmodel, type = "text",
+          title = "Summary of High Income Model",
+          align = TRUE,
+          out = "model_summary.txt")
+```
 
 # Results
+
+Labor Force Participation (lf_par_rate):
+Low-Income: Significant negative relationship.
+High-Income: Significant positive relationship.
+Insight: This contrast may indicate that in low-income countries, increases in labor force participation may include low-quality, low-income jobs, whereas in high-income countries, higher participation could be linked to high-skill, high-income roles.
+Trade as Percentage of GDP (trade_per):
+Both: Positive impact and significant in both models.
+Insight: Trade openness appears beneficial across both income groups, enhancing GNI per capita.
+Other Significant Predictors
+GDP Growth (gdp_growth):
+Low-Income: Insignificant.
+High-Income: Significant and positive.
+Insight: Economic growth plays a more critical role in high-income countries, possibly reflecting more developed financial systems and infrastructure which better translate growth into GNI per capita.
+Unemployment Rate (unem_rate):
+Low-Income: Insignificant.
+High-Income: Significant and negative.
+Insight: Unemployment negatively impacts GNI per capita in high-income countries, where unemployment may lead to significant economic losses.
+Inflation (inflation):
+Low-Income: Insignificant.
+High-Income: Significant and negative.
+Insight: Inflation is a critical factor in high-income countries, perhaps due to more developed financial markets and economies where inflation can erode real income levels more noticeably.
+Model Performance Metrics
+Log Likelihood and AIC:
+The log likelihood is notably lower (more negative) in the high-income model, reflecting the larger scale and complexity of the dataset.
+The AIC is also significantly higher, suggesting a less efficient model relative to the number of parameters and the sample size compared to the low-income model.
+Overall Conclusion
+The comparison reveals distinct economic dynamics between low- and high-income countries.
+Factors like labor force participation and trade are important in both, but their impacts differ by income level.
+Economic growth, unemployment, and inflation are more influential in high-income contexts, suggesting different policy priorities for each group.
+This analysis underscores the importance of tailored economic policies that consider specific national circumstances and economic structures.
 

### 2024-04-16 paper/paper.qmd
Writing model Latex and update result reference

@@
 \begin{align}
-\text{gni\_per\_cap} &= \alpha + \beta_1 \times \text{population}_i + \beta_2 \times \text{unem\_rate}_i + \beta_3 \times \text{lf\_par\_rate}_i \\
-&\quad + \beta_4 \times \text{trade\_per}_i + \beta_5 \times \text{inflation}_i, \\
+\text{gni\_per\_cap} &= \alpha + \beta_1 \times \text{gdp\_growth}_i + \beta_2 \times \text{population}_i + \beta_3 \times \text{unem\_rate}_i + \beta_4 \times \text{lf\_par\_rate}_i \\
+&\quad + \beta_5 \times \text{trade\_per}_i + \beta_6 \times \text{inflation}_i, \\
 \alpha &\sim \text{Normal}(0, 2.5), \\
@@
 As Economic growth rates and population can vary widely, and the impact on GNI per capita might scale logarithmically rather than linearly.
+
+\begin{align}
+\log(\text{gni\_per\_cap}) &= \alpha + \beta_1 \times \log(\text{gdp\_growth}_i+1) + \beta_2 \times \log(\text{population}_i +1) \\ 
+&\quad + \beta_3 \times \text{unem\_rate}_i + \beta_4 \times \text{lf\_par\_rate}_i \\
+&\quad + \beta_5 \times \text{trade\_per}_i + \beta_6 \times \text{inflation}_i, \\
+\alpha &\sim \text{Normal}(0, 2.5), \\
+\beta_1 &\sim \text{Normal}(0, 2.5), \\
+\beta_2 &\sim \text{Normal}(0, 2.5), \\
+\beta_3 &\sim \text{Normal}(0, 2.5), \\
+\beta_4 &\sim \text{Normal}(0, 2.5), \\
+\beta_5 &\sim \text{Normal}(0, 2.5)
+\end{align}
 
@@
 
-Labor Force Participation (lf_par_rate):
-Low-Income: Significant negative relationship.
-High-Income: Significant positive relationship.
-Insight: This contrast may indicate that in low-income countries, increases in labor force participation may include low-quality, low-income jobs, whereas in high-income countries, higher participation could be linked to high-skill, high-income roles.
-Trade as Percentage of GDP (trade_per):
-Both: Positive impact and significant in both models.
-Insight: Trade openness appears beneficial across both income groups, enhancing GNI per capita.
-Other Significant Predictors
-GDP Growth (gdp_growth):
-Low-Income: Insignificant.
-High-Income: Significant and positive.
-Insight: Economic growth plays a more critical role in high-income countries, possibly reflecting more developed financial systems and infrastructure which better translate growth into GNI per capita.
-Unemployment Rate (unem_rate):
-Low-Income: Insignificant.
-High-Income: Significant and negative.
-Insight: Unemployment negatively impacts GNI per capita in high-income countries, where unemployment may lead to significant economic losses.
-Inflation (inflation):
-Low-Income: Insignificant.
-High-Income: Significant and negative.
-Insight: Inflation is a critical factor in high-income countries, perhaps due to more developed financial markets and economies where inflation can erode real income levels more noticeably.
-Model Performance Metrics
-Log Likelihood and AIC:
-The log likelihood is notably lower (more negative) in the high-income model, reflecting the larger scale and complexity of the dataset.
-The AIC is also significantly higher, suggesting a less efficient model relative to the number of parameters and the sample size compared to the low-income model.
-Overall Conclusion
-The comparison reveals distinct economic dynamics between low- and high-income countries.
-Factors like labor force participation and trade are important in both, but their impacts differ by income level.
-Economic growth, unemployment, and inflation are more influential in high-income contexts, suggesting different policy priorities for each group.
-This analysis underscores the importance of tailored economic policies that consider specific national circumstances and economic structures.
+From the model we could see that in low-income countries the labour force participation rate has a significant negative value.
+Which in the high-income countries, the labour force participation rate has a significant positive value.
+This contrast may indicate that in low-income countries, increases in labor force participation may include low-quality, low-income jobs.
+Which are more unsecured during the COVID-19 crisis.
+Whereas in high-income countries, higher participation could be linked to more high-skill roles.
+[@job]
+The Trade as Percentage of GDP has positive impact and significant in both low and high income countries.
+Although the trade decreases sharply in 2020, however, the recovery in 2021 was more significant than the dropped in 2020.
+Causing the  total trade flows are above the pre-pandemic levels.[@Trade]
+
+The GDP Growth played an important role in the high-income countries and it has a significant and positive relationship.
+Possibly reflecting more developed financial systems and infrastructure which cause more growth into GNI per capita.
+However, it has insignificant negative relationship in the low-income coutries.
+
+The unemployment rate and the inflation rate both have a significant and negative impact in the high-income countries. 
+As the unemployment rate negatively impacts GNI per capita in high-income countries, where unemployment may lead to significant economic losses.
+The inflation is an important factor in high-income countries, however, due to more developed financial markets and economies, there are more ways to reduce the impact of inflation.
+Which in the low-income economies, the unemployment rate and the inflation rate have insignificant negative impact on GNI per capita.
 

### 2024-04-16 paper/paper.qmd
Working on discussion

@@
   - Sean Liu 
-thanks: "Code and data are available at: https://github.com/Sluuu/Final-Paper.git"
+thanks: "Code and data are available at: https://github.com/Sluuu/crisis-affect-countries.git"
 date: today
@@
 
-### Model Adjustment
-
+## Model Adjustment
 For a better performance and interpretation of the model, I have applied the log transformation to my model.
@@
 As Economic growth rates and population can vary widely, and the impact on GNI per capita might scale logarithmically rather than linearly.
+The following formula is the updated log transformation on certain variables.
 
@@
 
+\newpage
 ```{r}
@@
 
-## First discussion point {#sec-first-point}
+The results showed that the unemployment and inflation have significant negative impacts in high-income countries, as indicated by the strong and significant coefficients.
+These outcomes were unexpected, as the expectation was that these factors would have a larger negative effect in low-income countries than the high-income countries.
+It may be that high-income countries have a more developed financial markets and dependencies on stable economic indicators which are more sensitive to the fluctuations.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
+The negative coefficient for GDP growth in low-income countries was expected however, it was not significant.
+For high-income countries, GDP growth shows a positive and significant impact, indicating robust economic dynamics where growth directly enhances national income levels.
+Also the high income countries showed a significant stronger positive relationship on labour force participation rate and trade as a share of GDP.
 
@@
 
-If two different time events occurs does low-income countries being impacted differently before, after they become high-income countries?
+There are several weaknesses of this study, which in the low-income countries model several important variables like GDP growth, unemployment, and inflation did not show statistically significant effects.
+This might be due to the lack of sample sizes and the high fluctuations in the data.
+The second weakness was the lack of variables, the model did not include some relevant variables that affect GNI per capita, such as political stability, technology, and education levels, etc.
+Which both results might create biases onto the results.
+The next step of the study might be on increase the data size, such as include more years or some micro-level data.
 
@@
 
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
-
 \newpage

### 2024-04-16 paper/paper.qmd
Fixing typos

@@
 \text{gni\_per\_cap} &= \alpha + \beta_1 \times \text{gdp\_growth}_i + \beta_2 \times \text{population}_i + \beta_3 \times \text{unem\_rate}_i + \beta_4 \times \text{lf\_par\_rate}_i \\
-&\quad + \beta_5 \times \text{trade\_per}_i + \beta_6 \times \text{inflation}_i, \\
-\alpha &\sim \text{Normal}(0, 2.5), \\
-\beta_1 &\sim \text{Normal}(0, 2.5), \\
-\beta_2 &\sim \text{Normal}(0, 2.5), \\
-\beta_3 &\sim \text{Normal}(0, 2.5), \\
-\beta_4 &\sim \text{Normal}(0, 2.5), \\
-\beta_5 &\sim \text{Normal}(0, 2.5)
+&\quad + \beta_5 \times \text{trade\_per}_i + \beta_6 \times \text{inflation}_i
 \end{align}
@@
 &\quad + \beta_3 \times \text{unem\_rate}_i + \beta_4 \times \text{lf\_par\_rate}_i \\
-&\quad + \beta_5 \times \text{trade\_per}_i + \beta_6 \times \text{inflation}_i, \\
-\alpha &\sim \text{Normal}(0, 2.5), \\
-\beta_1 &\sim \text{Normal}(0, 2.5), \\
-\beta_2 &\sim \text{Normal}(0, 2.5), \\
-\beta_3 &\sim \text{Normal}(0, 2.5), \\
-\beta_4 &\sim \text{Normal}(0, 2.5), \\
-\beta_5 &\sim \text{Normal}(0, 2.5)
+&\quad + \beta_5 \times \text{trade\_per}_i + \beta_6 \times \text{inflation}_i
 \end{align}

### 2024-04-17 paper/paper.qmd
fixing format
Working on appendix

@@
 
-Lastly, the data presented in @fig-inflation, @fig-gdp-growth, and @fig-lf, I observed that high-income countries shared similar economic performance trends.
+Lastly, the data presented in @fig-inflation, @fig-gdp-growth, and @fig-ur, I observed that high-income countries shared similar economic performance trends.
 Despite not necessarily excelling in all economic indicators than the low-income countries, however, these countries generally maintain a more stable economic condition, with only a few exceptions or outliers noted.
+Which I will be expecting the data to be normally distributed.
 Also the overall economic performance in 2021 are better than in 2020.
@@
 ```
+
 The high income-countries have an overall better performance than the low-income countries: 
@@
 Also the high income countries showed a significant stronger positive relationship on labour force participation rate and trade as a share of GDP.
+From @fig-qq-low we could see that the observations were mostly separated and with multiple outliers which indicates that we have a low sample size also deviates from normality.
+Which shows that for the low-income countries model, it has potentially leading to skewed or biased estimates.
+From @fig-qq-high the points on the Q-Q plot fall approximately along a straight line, suggesting that the residuals follow a normal distribution.
+Which it also has less outliers compared with the low-income model.
+It also furthe confirm that the residuals are normally distributed, supporting the validity of the least squares regression assumptions.
 
@@
 The next step of the study might be on increase the data size, such as include more years or some micro-level data.
-
-\newpage
-
-\appendix
 
@@
 
+```{r}
+#| label: fig-qq-low 
+#| fig-cap: Q-Q Plot of the low-income model
+#| echo: false
+#| warning: false
+#| message: false
+
+# Q-Q Plot
+qqnorm(residuals(log_lowmodel), main = "Q-Q Plot of the low-income model")
+
+```
+
+```{r}
+#| label: fig-qq-high
+#| fig-cap: Q-Q Plot of the high-income model
+#| echo: false
+#| warning: false
+#| message: false
+
+# Q-Q Plot
+qqnorm(residuals(log_highmodel), main = "Q-Q Plot of the high-income model")
+```
+
 \newpage

### 2024-04-17 scripts/04-model.R
Save Models

@@ -1,37 +1,56 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Modes for low and high income countries
+# Author: Sean Liu
+# Date: 16 April 2024
+# Contact: yuhsiang.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(readr)
+library(dplyr)
+library(arrow)
+library(dataverse)
+library(WDI)
 library(rstanarm)
+library(broom)
+library(stargazer)
 
 #### Read data ####
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+lowincome <- data_na %>% filter(low_income == 1)
+highincome <- data_na %>% filter(low_income == 0)
+lowmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 
+                family = gaussian(link = "identity"), data = lowincome)
+
+stargazer(lowmodel, type = "text",
+          title = "Summary of Low Income Model",
+          align = TRUE,
+          out = "model_summary.txt")
+
+
+highmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 
+                family = gaussian(link = "identity"), data = highincome)
+
+stargazer(lowmodel, type = "text",
+          title = "Summary of High Income Model",
+          align = TRUE,
+          out = "model_summary.txt")
 
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  lowmodel,
[…]

### 2024-04-17 scripts/04-model.R
Working on model

@@ -19,8 +19,15 @@ library(stargazer)
 
 #### Read data ####
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+data_na<- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+
 
 ### Model data ####
+
+data_na <- data_na %>%
+  mutate(low_income = as.numeric(gni_per_cap <= 3995))
+
+
 lowincome <- data_na %>% filter(low_income == 1)
 highincome <- data_na %>% filter(low_income == 0)
 lowmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 

### 2024-04-17 scripts/04-model.R
Fix typo

@@ -18,7 +18,6 @@ library(broom)
 library(stargazer)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 data_na<- read_csv(here::here("data/analysis_data/analysis_data.csv"))
 
 

### 2024-04-17 scripts/02-data_cleaning.R
Fixing typo

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw data, filtering by year 2008, 2009, and 2020
+# Purpose: Cleans the raw data, filtering by year 2020 and 2021
 # Author: Sean Liu
 # Date: 2 April 2024
 # Contact: yuhsiang.liu@mail.utoronto.ca

### 2024-04-17 paper/paper.qmd
fixing from feedabck

@@
 A country that has a GNI per capita larger than \$3995 will be categorized as high-income country.
-Which I have grouped out in 2020 there are 27 low-income countries in this definition and 85 high-income countries.
-In 2021, there are
+Which I have grouped out in 2020 there are 27 low-income countries and 85 high-income countries in this definition.
+In 2021, there are 26 low-income countries and 85 high-income countries.
 

### 2024-04-18 paper/paper.qmd
Fixing grammer

@@
 abstract: "This paper examines the impacts of global crises focusing on the 2020 and 2021 COVID-19 pandemic's effects on low and high-income countries.
-It highlights how globalization links economies, leading to shared challenges during worldwide disturbances.
-The analysis reveals that low-income countries have suffered more acutely, experiencing higher unemployment, inflation, and slower GDP growth."
+It highlights how globalization links economies, leading to shared challenges during the worldwide pandemic.
+The analysis shows that low-income countries have suffered more than high-income countries, by experiencing higher unemployment, inflation, and slower GDP growth."
 format: pdf
@@
 
-In this paper, my focus is on the varying impacts of global crises on low and high income countries.
+In this paper, my focus is on the different impacts of global crises on low and high income countries.
 I have selected the 2020 and 2021 COVID-19 pandemic as the primary subject of my study.
-This crises have led to varying degrees of economic uncertainty and fluctuations in economic performance worldwide.[@compare] The World Bank indicates that the COVID-19 pandemic has widen the inequality between low and high-income countries, with low-income countries experiencing higher unemployment and inflation rates, along with slower GDP growth.
+This crises have led to varying degrees of economic uncertainty and fluctuations in economic performance worldwide.[@compare] The World Bank indicates that the COVID-19 pandemic has widen the inequality between low and high-income countries, with low-income countries experiencing higher unemployment and inflation rates, also with slower GDP growth.
 These trends align with the expectations set in my study.[@inequality]
@@
 Lastly, the data presented in @fig-inflation, @fig-gdp-growth, and @fig-ur, I observed that high-income countries shared similar economic performance trends.
-Despite not necessarily excelling in all economic indicators than the low-income countries, however, these countries generally maintain a more stable economic condition, with only a few exceptions or outliers noted.
-Which I will be expecting the data to be normally distributed.
+Although not necessarily always better in all economic indicators than the low-income countries, however, these countries generally maintain a more stable economic condition, with only a few exceptions or outliers noted.
+Therefore, I will be expecting the data to be normally distributed.
 Also the overall economic performance in 2021 are better than in 2020.
@@
 I will explore how different variables affect `gni_per_cap` in both low and high-income countries.
-Specifically, I anticipate that low-income countries will experience larger negative effect from unemployment and inflation rates, as well as experience negative GDP growth.
-This analysis aims to provide insights into the economic challenges faced by these countries and inform potential policy interventions.
+Specifically, I expect that low-income countries will experience larger negative effect from unemployment and inflation rates, as well as experience negative or slower GDP growth.
+This analysis aims to provide insights into the economic challenges faced by these countries and inform potential policy impacts.
 
@@
 The Labor force participation rate (`lf_par_rate`) coefficient is -52.818 with a standard error of 13.104, and it is significant at the p<0.01 level.
-This significant negative relationship implies that higher labor force participation is associated with lower GNI per capita among the low-income countries in this analysis, which might suggest underemployment or low-quality jobs prevalent in these economies.
+This significant negative relationship shows that higher labor force participation is connected with lower GNI per capita among the low-income countries in this analysis, which might suggest underemployment or low-quality jobs in these economies.
 The Trade percentage of GDP (`trade_per`) coefficient is 11.697 with a standard error of 3.992, significant at the p<0.01 level.
@@
 The Trade Percentage of GDP (`trade_per`) coefficient is 59.875 with a standard error of 22.554, significant at the p<0.01 level.
-This indicates a robust positive relationship, suggesting that trade openness contributes significantly to GNI per capita in high-income countries.
+This shows a robust positive relationship, suggesting that trading contributes significantly to GNI per capita in high-income countries.
 Lastly, the Inflation (`inflation`) coefficient is -2,799.239 with a standard error of 600.123, significant at the p<0.01 level.
-This suggests a strong negative impact of inflation on GNI per capita, indicating that higher inflation rates are detrimental to income levels.
+This suggests a strong negative impact of inflation on GNI per capita, indicating that higher inflation rates are fluctuating income levels.
 
@@
 The log of `gdp_growth` and `population` shared similar reasons.
-As Economic growth rates and population can vary widely, and the impact on GNI per capita might scale logarithmically rather than linearly.
+As economic growth rates and population can vary widely, and the impact on GNI per capita might scale logarithmically other than linearly.
 The following formula is the updated log transformation on certain variables.
@@
 Which are more unsecured during the COVID-19 crisis.
-Whereas in high-income countries, higher participation could be linked to more high-skill roles.
+Whereas in high-income countries, higher participation could be linked to more high-skilled roles.
 [@job]
@@
 Possibly reflecting more developed financial systems and infrastructure which cause more growth into GNI per capita.
-However, it has insignificant negative relationship in the low-income coutries.
+However, it has insignificant negative relationship in the low-income countries.
 
@@
 The negative coefficient for GDP growth in low-income countries was expected however, it was not significant.
-For high-income countries, GDP growth shows a positive and significant impact, indicating robust economic dynamics where growth directly enhances national income levels.
+For high-income countries, GDP growth shows a positive and significant impact, indicating strong economic dynamics and growth directly improving national income levels.
 Also the high income countries showed a significant stronger positive relationship on labour force participation rate and trade as a share of GDP.
 From @fig-qq-low we could see that the observations were mostly separated and with multiple outliers which indicates that we have a low sample size also deviates from normality.
-Which shows that for the low-income countries model, it has potentially leading to skewed or biased estimates.
+Which shows that for the low-income countries model, it has a probability of leading to skewed or biased estimates.
 From @fig-qq-high the points on the Q-Q plot fall approximately along a straight line, suggesting that the residuals follow a normal distribution.
 Which it also has less outliers compared with the low-income model.
-It also furthe confirm that the residuals are normally distributed, supporting the validity of the least squares regression assumptions.
+It also further confirm that the residuals are normally distributed.
 
@@
 There are several weaknesses of this study, which in the low-income countries model several important variables like GDP growth, unemployment, and inflation did not show statistically significant effects.
-This might be due to the lack of sample sizes and the high fluctuations in the data.
+This might be due to the lack of sample sizes and the high variations in the data.
 The second weakness was the lack of variables, the model did not include some relevant variables that affect GNI per capita, such as political stability, technology, and education levels, etc.
